---
title: "Assignment 1 MVA"
output: html_document
date: "2024-02-11"
---
1. Ask an important question you want answered (1 at least) : 
Answer:


Question 1: How does the price of gold relate to various economic indicators and market factors?


Question 2: Variable Impact of economic indicators and market factors with respect to time (Year-Month trend)

2. Answer why this question(s) is/are important to you: 

Answer: 

Understanding the relationship between the price of gold and economic indicators can provide insights into market trends, investment opportunities, and potential risks. For example, knowing how gold prices respond to changes in interest rates, inflation, and currency exchange rates can help investors make informed decisions about portfolio allocation and risk management.


3. Find and collect data
(need one dependent variable and more than 3 dependent variables): 

Answer: 

Data Collection:

Source: Kaggle
1.	Dependent Variable (Target):
•    Gold Price
2.	Independent Variables:
•     Crude Oil Price
•     Interest Rate
•    USD/INR Exchange Rate
•    Sensex (Stock Market Index)
•    CPI (Consumer Price Index)
•    USD Index

4. Describe your data (create a data dictionary): 4 Points

Answer: 
Attribute/Variable description:

•	Gold_Price: The price of gold per unit (e.g., ounce) over time.
•	Crude_Oil: The price of crude oil per barrel over time.
•	Interest_Rate: The prevailing interest rate over time.
•	USD_INR: The exchange rate of US dollars to Indian rupees over time.
•	Sensex: The value of the Sensex (stock market index) over time.
•	CPI: The Consumer Price Index measuring inflation over time.
•	USD_Index: The value of the US Dollar Index representing the value of the US dollar relative to a basket of foreign currencies over time.


Dictionary created below
```{r}
library(readxl)
GoldUP <- read_excel("C:/Rutgers/Subjects/Spring Sem/Multivariate Analysis/GoldUP.xlsx")
View(GoldUP)
```

Data Dictionary with data
```{r}
# Create a data frame with variable names, descriptions, and data points
data_info <- data.frame(
  Variable_Name = c("Date","Gold_Price", "Crude_Oil", "Interest_Rate", "USD_INR", "Sensex", "CPI", "USD_Index"),
  Description = c("Date","The price of gold per unit (e.g., ounce) over time.",
                  "The price of crude oil per barrel over time.",
                  "The prevailing interest rate over time.",
                  "The exchange rate of US dollars to Indian rupees over time.",
                  "The value of the Sensex (stock market index) over time.",
                  "The Consumer Price Index measuring inflation over time.",
                  "The value of the US Dollar Index representing the value of the US dollar relative to a basket of foreign currencies over time."),
  Data_Points = sapply(GoldUP, function(x) paste(head(x, 5), collapse = ", "))
)

# Print the data frame
print(data_info)
```
```{r}
# Descriptive statistics
summary(GoldUP)
is.na(GoldUP)
sapply(GoldUP, mean)
sapply(GoldUP, sd)

# Check data types
str(GoldUP)

# Convert non-numeric variables to numeric
# Convert non-numeric variables to numeric
GoldUP$Gold_Price <- as.numeric(GoldUP$Date)

GoldUP$Gold_Price <- as.numeric(GoldUP$Gold_Price)
GoldUP$Crude_Oil <- as.numeric(GoldUP$Crude_Oil)
GoldUP$Interest_Rate <- as.numeric(GoldUP$Interest_Rate)
GoldUP$USD_INR <- as.numeric(GoldUP$USD_INR)
GoldUP$Sensex <- as.numeric(GoldUP$Sensex)
GoldUP$CPI <- as.numeric(GoldUP$CPI)
GoldUP$USD_Index <- as.numeric(GoldUP$USD_Index)
# Repeat this for all variables if needed

# Check data types again
str(GoldUP)

# Check for non-numeric values in each column
sapply(GoldUP, function(x) any(!is.na(as.numeric(x))))

# Convert non-numeric values to NA
GoldUP[] <- lapply(GoldUP, function(x) ifelse(is.na(as.numeric(x)), NA, as.numeric(x)))

# Compute correlations
correlation_matrix <- cor(GoldUP, use = "complete.obs")
print(correlation_matrix)


# Check the class of your data
class(GoldUP)

# If it's not a dataframe, convert it to a dataframe
if (!is.data.frame(GoldUP)) {
  GoldUP <- as.data.frame(GoldUP)
}

# Check the variable names in your dataframe
names(GoldUP)

# Make sure the variable names are referenced correctly in your model formula
# For example:
model <- lm(Gold_Price ~ Crude_Oil + Interest_Rate + USD_INR + Sensex + CPI + USD_Index, data = GoldUP)

summary(model)
```